---
title: "Working with the manifestoR package"
author: "Jirka Lewandowski <jirka.lewandowski@wzb.eu>"
date: "14/09/2015"
output:
  pdf_document:
    number_sections: yes
    toc: yes
  html_document:
    keep_md: yes
    number_sections: yes
    toc: yes
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{The workflow with the manifestoR package}
\usepackage[utf8]{inputenc}
-->

\newpage

# Downloading documents from the Manifesto Corpus

When publishing work using the Manifesto Corpus, please make sure to cite it correctly and to give the identification number of the corpus version used for your analysis.

You can print citation and version information with the function `mp_cite()`.

## Loading the package

First of all, load the `manifestoR` package with the usual R syntax:

```{r eval=FALSE}
library(manifestoR)
```

```{r eval=TRUE, include=FALSE}
library(manifestoR)
```

## Connecting to the Manifesto Project Database API

To access the data in the Manifesto Corpus, an account for the Manifesto Project
webpage with an API key is required.
If you do not yet have an account, you can create one at 
[https://manifesto-project.wzb.eu/signup]().
If you have an account, you can create and download the API key on your profile page.

For every R session using manifestoR and connecting to the Manifesto Corpus database,
you need to set the API key in your work environment.
This can be done by passing either a key or the name of a file containing the
key to manifestoR's `mp_setapikey()` function (see documentation `?mp_setapikey`
for details).
Thus, your R script using manifestoR usually will start like this:

```{r eval=FALSE}
library(manifestoR)
mp_setapikey("manifesto_apikey.txt")
```
```{r eval=TRUE, include=FALSE}
mp_setapikey(key.file = "../tests/manifesto_apikey.txt")
```

This R code presumes that you have stored and downloaded the API key in a file
name `manifesto_apikey.txt` in your current R working directory.

Note that it is a security risk to store the API key file or a script containing
the key in public repositories.

## Downloading the Manifesto Project Dataset

You can download the Manifesto Project Dataset (MPDS) with the function `mp_maindataset()`.
By default the most recent update is returned, but you can specify older versions
to get for reproducibility (type `mp_coreversions()` for a list of version and `?mp_maindataset`
for usage information). For analysing the dataset using scaling functions, refer
to the section [Using manifestoR's scaling functions](#using-manifestors-scaling-functions) below.

## Downloading documents

(Bulk-)Downloading documents works via the function `mp_corpus(...)`.
It can be called with a logical expression specifying the subset of the Manifesto
Corpus that you want to download:

```{r, warning=FALSE}
my_corpus <- mp_corpus(countryname == "Austria" & edate > as.Date("2000-01-01"))
my_corpus
```

`mp_corpus` returns a `ManifestoCorpus` object, a subclass of `Corpus` as defined
in the natural language processing package `tm` (Feinerer & Hornik 2015).
Following `tm`s logic, a `ManifestoCorpus` consists of `ManifestoDocuments`.
For both, corpus and documents, `tm` provides accessor functions to the corpus
and documents content and metadata:

```{r}
head(content(my_corpus[[1]]))
meta(my_corpus[[1]])
```

For more information on the available metadata per document, refer to the section
[Using the document metadata](#using-the-document-metadata) below.
For more information on how to use the text mining functions provided by `tm`
for the data from the Manifesto Corpus, refer to the section
[Processing and analysing the corpus documents](#processing-and-analysing-the-corpus-documents) below.

The variable names in the logical expression used for querying the corpus
database (`countryname` and `edate` in the example above) can be any column names from the
Manifesto Project's Main Dataset or your current R environment. The Main Dataset
itself is available in manifestoR via the funcion `mp_maindataset()`:

```{r, warning = FALSE}
mpds <- mp_maindataset()
print(head(names(mpds)))
mp_corpus(rile > 60) ## another example of data set based corpus query
```

Alternatively, you can download election programmes on an individual basis
by listing combinations of party ids and election dates in a `data.frame`
and passing it to `mp_corpus(...)`:

```{r warning=FALSE}
wanted <- data.frame(party=c(41220, 41320),
                     date=c(200909, 200909))
mp_corpus(wanted)
```

The party ids (41220 and 41320 in the example) are the ids as in the Manifesto Project's
main dataset.
They can be found in the current dataset documentation at
[https://manifesto-project.wzb.eu/datasets]() or in the main dataset.

Note that we received only 1 document, while querying for two.
This is because the party with the id 41220 (KPD) did not run for elections
in September 2009.
Also, not for every party and election data manifesto documents are availabe in the Manifesto Project Corpus.
You can check availability of your query beforehand with the function
`mp_availability(...)`:
```{r}
mp_availability(party == 41113)
```

Downloaded documents are automatically cached locally. To learn about
the caching mechanism read the section
[Efficiency and reproducibility: caching and versioning](#efficiency-and-reproducibility-caching-and-versioning)
below.

## Viewing original documents

Apart from the machine-readable, annotated documents, the Manifesto Corpus also
contains original layouted election programmes in PDF format. If available, they
can be viewed via the function `mp_view_originals(...)`, which takes exactly the
format of arguments as `mp_corpus(...)` ([see above](#downloading-documents)), e.g.:

```{r eval = FALSE}
mp_view_originals(party == 41320 & date == 200909)
```

The original documents are shown in you system's web browser. All URLs opened
by this function refer only to the Manifesto Project's Website. If you want to
open more than 5 PDF documents at once, you have to specify the maximum number
of URLs allows to be opened manually via the parameter `maxn`. Since opening
URLs in an external browser costs computing resources on your local machine,
make sure to use only values for `maxn` that do not slow down or make your computer
unresponsive.

```{r eval = FALSE}
mp_view_originals(party > 41000 & party < 41999, maxn = 20)
```


# Processing and analysing the corpus documents

As in `tm`, the textual content of a document is returned by the function `content`:

```{r}
txt <- content(my_corpus[[2]])
class(txt)
head(txt, n = 4)
```

## Working with the CMP codings

The central way for accessing the CMP codings is the accessor method `codes(...)`.
It can be called on `ManifestoDocument`s and `ManifestoCorpus`s and returns a vector
of the CMP codings attached to the quasi-sentences of the document/corpus in a row:

```{r}
doc <- my_corpus[[2]]
head(codes(doc), n = 15)
head(codes(my_corpus), n = 15)
```

Thus you can for example use R's functionality to count the codes or select quasi-
sentences (units of texts) based on their code:
```{r}
table(codes(doc))

doc_subcodes <- subset(doc, codes(doc) %in% c(202, 503, 607))
length(doc_subcodes)
length(doc_subcodes)/length(doc)
```

The CMP coding scheme can be found in the online documentation of the Manifesto
Project dataset at [https://manifesto-project.wzb.eu/coding_schemes/1]().

## Working with additional layers of codings

Besides the main layer of CMP codings, you can create, store and access additional
layers of codings in `ManifestoDocument`s by passing a name of the coding layer 
as additional argument to the function `codes()`:

```{r}
## assigning a dummy code of alternating As and Bs
codes(doc, "my_code") <- rep_len(c("A", "B"), length.out = length(doc))
head(codes(doc, "my_code"))
```

You can view the names of the coding layers stored in a `ManifestoDocument` with
the function `code_layers()`:
```{r}
code_layers(doc)
```

Note that certain documents downloaded from the Manifesto Corpus Database already
have a second layer of codes named `eu_code`. These are codes that have been assigned
to quasi-sentences by CMP coders additionally to the main CMP code to indicate
policy statements that should or should not be implemented on the level of the
European union. The documents that were coded in this way are marked in the
corpus' metadata with the flag `has_eu_code`
(see below [Using the document metadata](#using-the-document-metadata)).
Note that, since these codes also have been used for computing the `per` and `rile`
variables in the Manifesto Project Main Dataset, they are also used in `manifestoR`s
`count_codes` and `rile` functions (see below [Scaling texts](#scaling-texts)) if the respective metadata flag is present.

## Text mining tools

Since the Manifesto Corpus uses the infrastructure of the `tm` package
(Feinerer & Hornik 2015), all of `tm`s filtering and transformation functionality
can be applied directly to the downloaded `ManifestoCorpus`.

For example, standard natural language processors are available to clean the corpus:

```{r}
head(content(my_corpus[[3]]))
corpus_cleaned <- tm_map(my_corpus, removePunctuation)
corpus_nostop <- tm_map(corpus_cleaned, removeWords, stopwords("german"))
head(content(corpus_nostop[[3]]))
```

So is analysis in form of term document matrices:
```{r}
tdm <- TermDocumentMatrix(corpus_nostop)
inspect(tdm[c("menschen", "wahl", "familie"),])
findAssocs(tdm, "stadt", 0.97) ## find correlated terms, see ?tm::findAssocs
```

For more information about the functionality provided by the `tm`,
please refer to its [documentation](http://cran.r-project.org/web/packages/tm/vignettes/tm.pdf).

## Selecting relevant parts of text

For applications in which not the entire text of a document is of interest, but
rather a subset of the quasi-sentences matching certain criteria,
`manifestoR` provides a function `subset(...)` working just like R's internal
`subset` function.

It can, for example, be used to filter quasi-sentences based on codes or the text:

```{r}
# subsetting based on codes (as example above)
doc_subcodes <- subset(doc, codes(doc) %in% c(202, 503, 607))
length(doc_subcodes)

# subsetting based on text
doc_subtext <- subset(doc, grepl("Demokratie", content(doc)))
head(content(doc_subtext), n = 3)
head(codes(doc_subtext), n = 10)
```

Via `tm_map` the filtering operations can also be applied to an entire corpus:
```{r}
corp_sub <- tm_map(my_corpus, function(doc) {
  subset(doc, codes(doc) %in% c(202, 503, 607))
})
head(content(corp_sub[[3]]))
head(codes(corp_sub))
```

For convenience, it is also possible to filter quasi-sentences with specific
codes directly when downloading a corpus. For this, the additional argument
`codefilter` with a list of CMP codes of interest is passed to `mp_corpus`:
```{r, warning=FALSE}
corp_sub <- mp_corpus(countryname == "Australia", codefilter = c(103, 104))
head(content(corp_sub[[1]]))
head(codes(corp_sub))
```

## Using the document metadata

Each document in the Manifesto Corpus has meta information about itself attached.
They can be accessed via the function `meta`:
```{r}
meta(doc)
```

It is possible to access and also modify specific metadata entries:
```{r}
meta(doc, "party")
meta(doc, "manual_edits") <- TRUE
meta(doc)
```

Document metadata can also be bulk-downloaded with the function `mp_metadata`,
taking the same set of parameters as `mp_corpus`:
```{r warning=FALSE}
metas <- mp_metadata(countryname == "Spain")
head(metas)
```

The field ...

* ... `party` contains the party id from the Manifesto Project Dataset.
* ... `date` contains the month of the election in the same format as in the
Manifesto Project Dataset (`YYYYMM`)
* ... `language` specifies the language of the document as a word.
* ... `is_primary_doc` is FALSE only in cases where for a single party and
election date multiple manifestos are available and this is the document not used
for coding by the Manifesto Project.
* ... `may_contradict_core_dataset` is TRUE for documents where the CMP codings
in the corpus documents might be inconsistent with the coding aggregates in the
Manifesto Project's Main Dataset. This applies to manifestos which have been either
recoded after they entered the dataset or cases where the dataset entries are
derived from hand-written coding sheets used prior to the digitalization of the
Manifesto Project's data workflow, but the documents were digitalized and added
to the Manifesto Corpus afterwards.
* ... `annotations` is TRUE whenenver there are CMP codings available for the document.
* ... `has_eu_code` marks document in which the additional coding layer `eu_code` is present. These codes have been assigned to quasi-sentences by CMP coders additionally to the main CMP code to indicate policy statements that should or should not be implemented on the level of the European union.

The other metadata entries have primarily technical functions for communication
between the `manifestoR` package and the online database.


# Efficiency and reproducibility: caching and versioning

To save time and network traffic, `manifestoR` caches all downloaded data and
documents in your computer's working memory and connects to the online database
only when data is required that has not been downloaded before.

```{r eval=TRUE, include=FALSE}
mp_emptycache()
```

```{r message=TRUE, warning=FALSE}
corpus <- mp_corpus(wanted)
subcorpus <- mp_corpus(wanted[3:7,])
```

Note that in the second query no message informing about the connection to the
Manifesto Project's Database is printed, since no data is actually downloaded.

This mechanism also ensures **reproducibility** of your scripts, analyses
and results: executing your code again will yield the same results, even if
the Manifesto Project's Database is updated in the meantime.
Since the cache is only stored in the working memory, however, in order to ensure
reproducibility across R sessions, it is advisable to
**save the cache to the hard drive** at the end of analyses and load it in the
beginning:

```{r eval=FALSE}
mp_save_cache(file = "manifesto_cache.RData")

## ... start new R session ... then:

library(manifestoR)
mp_setapikey("manifesto_apikey.txt")
mp_load_cache(file = "manifesto_cache.RData")
```

This way `manifestoR` always works with the same snapshot of the Manifesto Project
Database and Corpus, saves a lot of unnecessary online traffic and also enables
you to continue with your analyses offline.

Each snapshot of the Manifesto Corpus is identified via a version number, which is
stored in the cache together with the data and can be accessed via
```{r}
mp_which_corpus_version()
```

When collaborating on a project with other researchers, it is advisable to use
the same corpus version for reproducibility of the results.
`manifestoR` can be set to use a specific version with the functions
```{r}
mp_use_corpus_version("2015-3")
```

In order to guarantee reproducibility of **published work**, please also mention
the corpus version id used for the reported analyses in the publication.

For updating locally cached data to the most recent version of the
Manifesto Project Corpus, `manifestoR` provides two functions:

```{r}
mp_check_for_corpus_update()
mp_update_cache()
mp_check_for_corpus_update()
```

For more detailed information on the caching mechanism and on how to use and load
specific snapshots of the Manifesto Corpus, refer to the R documentations of the
functions mentioned here as well `mp_use_corpus_version`, `mp_corpusversions`,
`mp_which_corpus_version`.


# Exporting documents

If required `ManifestoCorpus` as well as `ManifestoDocument` objects can be
converted to R's internal `data.frame` format and processed further:
```{r}
doc_df <- as.data.frame(doc)
head(within(doc_df, {
  ## for pretty printing
  text <- paste0(substr(text, 1, 60), "...")
}))
```

The function also provides a parameter to include all available metadata in
the export:
```{r}
doc_df_with_meta <- as.data.frame(doc, with.meta = TRUE)
print(names(doc_df_with_meta))
```

For more information on the available metadata per document, refer to the section
[Using the document metadata](#using-the-document-metadata) above.

Note again that also all functionality provided by `tm`, such as `writeCorpus`
is available on a `ManifestoCorpus`.

# Scaling texts

Scaling of political content refers to the estimation of its location in a policy
space (Grimmer & Stewart 2013). `manifestoR` provides several functions to scale coded documents by
known routines such as the RILE measure (see sections
[Using manifestoR's scaling functions](#using-manifestors-scaling-functions)),
as well as infrastructure to create new scales (see section [Writing custom scaling functions](#writing-custom-scaling-functions)) and statistical analysis routines for the
distributions of scaling functions (see section [Bootstrapping scaling function distributions and standard errors](#bootstrapping-scaling-function-distributions-and-standard-errors)).

## Using manifestoR's scaling functions

Implementationwise, a scaling function in `manifestoR` takes a data.frame of cases
and outputs a position value for each case. The Manifesto Project Dataset (MPDS)
can be downloaded in `manifestoR` using the function `mp_maindataset()` (see section
[Downloading the Manifesto Project Dataset](#downloading-the-manifesto-project-dataset) above). Then you can e.g. compute the RILE scores of cases from the main dataset by calling:

```{r}
mpds <- mp_maindataset()
rile(subset(mpds, countryname == "Albania"))
```

What variables are used from the input data depends on the scaling function. All
currently implemented functions use only the percentages of coded categories,
in the form of variables starting with "per" as in the Manifesto Project Dataset.
The following functions are currently available:

* RILE according to Laver & Budge (1992): `rile`
* logit rile according to Lowe et al. (2011): `logit_rile`
* Vanilla scaling according to Gabel & Huber (2000): `vanilla`
* ... (more scaling functions are planned and contributions are welcome, see
[Contributing to manifestoR](#contributing-to-manifestor)).

To apply scaling functions directly to coded documents or corpora
you can use the function `mp_scale`. It takes a `ManifestoCorpus`
or `ManifestoDocument` and returns the scaled positions for each document:

```{r message=FALSE}
corpus <- mp_corpus(countryname == "Romania")
mp_scale(corpus, scalingfun = logit_rile)
```

## Writing custom scaling functions

Writing custom scaling functions for texts in `manifestoR` is easy, since it
requires nothing more than writing a function that takes a `data.frame` of cases
as input and returns a vector of values. `mp_scale` provides the mechanism
that converts a coded text to a `data.frame` with "per" variables such that your function
can handle it:

```{r}
custom_scale <- function(data) { 
  data$per402 - data$per401
}
mp_scale(corpus, scalingfun = custom_scale)
```

In addition, `manifestoR` provides several function templates you can use for
creating scales, e.g. a weighted sum of per variables
(`scale_weighted`), the logit ratio of category counts (`scale_logit`)
or ratio scaling as suggested by Kim and Fording (1998) and by Laver & Garry (2000)
(`scale_ratio`). For example, the ratio equivalent to the simple function
above can be implemented as:
```{r}
custom_scale <- function(data) {
    scale_ratio(data, pos = c("per402"), neg = c("per401"))
}
mp_scale(corpus, scalingfun = custom_scale)
```

For details on these template functions, their parameters and how to use them,
see the R documentation `?scale`.

## Bootstrapping scaling function distributions and standard errors

In order to better evaluate the significance of analysis results based on
scaled coded texts, Benoit, Mikhaylov, and Laver (2009) proposed to approximate
the standard errors of the scale variable by bootstrapping its distribution.
This procedure is available via the function `mp_bootstrap`:

```{r}
data <- subset(mpds, countryname == "Albania")
mp_bootstrap(data, fun = rile)
```

Note that the argument `fun` can be any scaling function and the returned `data.frame`
containes the scaled position as well as the bootstrapped standard deviations.
Also, with the additional parameters `statistics`, you can compute arbitrary statistics
from the bootstrapped distribution, such as variance or quantiles:
```{r}
custom_scale <- function(data) { 
  data$per402 - data$per401
}
mp_bootstrap(data,
             fun = custom_scale,
             statistics = list(var, 0.025, 0.975))
```


# Additional Information

For a more detailed reference and complete list of the functions provided
by `manifestoR`, see the R package reference manual on CRAN:
[http://cran.r-project.org/web/packages/manifestoR/manifestoR.pdf](http://cran.r-project.org/web/packages/manifestoR/manifestoR.pdf)

## Contacting the Manifesto Project team

You can get in touch with the Manifesto Project team by e-mailing to
[manifesto-communication@wzb.eu](mailto:manifesto-communication@wzb.eu).
We are happy to receive your feedback and answer questions about the Manifesto
Corpus, including errors or obscurities in the corpus documents. In this case
please make sure to include the party id, election date and the corpus version
you were working with (accessible via `mp_which_corpus_version`).
For general questions about the Project and dataset, please check the
[Frequently Asked Questions](https://manifesto-project.wzb.eu/questions) section
on our website first.

## Contributing to manifestoR

We welcome bug reports, feature requests or (planned) source code contributions for the
`manifestoR` package. For all of these, best refer to our repository on github:
[https://github.com/ManifestoProject/manifestoR](https://github.com/ManifestoProject/manifestoR).
For more information, please refer to the Section "Developing" in the README file
of the github repository.

\newpage

# References

Benoit, K., Laver, M., & Mikhaylov, S. (2009). Treating Words as Data with Error: Uncertainty in Text Statements of Policy Positions. American Journal of Political Science, 53(2), 495-513. http://doi.org/10.1111/j.1540-5907.2009.00383.x

Feinerer, I., & Hornik, K. (2015). Tm: Text Mining Package. http://cran.r-project.org/web/packages/tm/index.html

Gabel, M. J., & Huber, J. D. (2000). Putting Parties in Their Place: Inferring Party Left-Right Ideological Positions from Party Manifestos Data. American Journal of Political Science, 44(1), 94-103.

Grimmer, J., & Stewart, B.. 2013. Text as Data: The Promise and Pitfalls of Automatic Content Analysis Methods for Political Texts. Political Analysis 21(3): 267–97.

Kim, H., & Fording, R. C. (1998). Voter ideology in western democracies, 1946-1989. European Journal of Political Research, 33(1), 73-97.

Laver, M. & Budge, I., eds. (1992). Party Policy and Government Coalitions, Houndmills, Basingstoke, Hampshire: The MacMillan Press 1992

Laver, M., & Garry, J. (2000). Estimating Policy Positions from Political Texts. American Journal of Political Science, 44(3), 619-634.

Lehmann, P.,  Matthieß, T., Merz, N., Regel, S., & Werner, A. (2015): Manifesto Corpus. Version: 2015-4. Berlin: WZB Berlin Social Science Center.

Lowe, W., Benoit, K., Mikhaylov, S., & Laver, M. (2011). Scaling Policy Preferences from Coded Political Texts. Legislative Studies Quarterly, 36(1), 123-155.

Volkens, A., Lehmann, P., Matthieß, T., Merz, N., Regel, S., & Werner, A (2015): The Manifesto Data Collection. Manifesto Project (MRG/CMP/MARPOR). Version 2015a. Berlin: Wissenschaftszentrum Berlin für Sozialforschung (WZB)

